---
title: "_TestCode"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

# Prep Code

```{r prep echo=TRUE include=FALSE}
library(here)
here()

source(here::here("_code/_common.R"))

selection_table_path <-  "E:/Rails/recordings/test/SPP-DD10_48c_20250716/birdnetLocal_SPP-DD10_48c_20250716/BirdNET_SelectionTable_SPP-DD10_48c_20250716.txt"

test <- process_birdnet_table(selection_table_path)

```

```{r old}

myBirdnetFile <-  "E:/Rails/recordings/test/SPP-DD10_48c_20250716/birdnetLocal_SPP-DD10_48c_20250716/Appended_BirdNET_SelectionTable_SPP-DD10_48c_20250716.txt"

# Read selection table
myBirdnet <- read_delim(myBirdnetFile, delim = "\t")%>%
  clean_names()

first <- "E:/Rails/recordings/TJE-01/TJE-01_48c_20250515/birdnetLocal_TJE-01_48c_20250515/BirdNET_SelectionTable_TJE-01_48c_20250515.txt"
second <- "E:/Rails/recordings/TJE-01/TJE-01_48c_20250515/birdnetLocal_TJE-01_48c_20250515/Appended_BirdNET_SelectionTable_TJE-01_48c_20250515.txt"

# Read selection table
table1 <- read_delim(first, delim = "\t")
table2 <- read_delim(second, delim = "\t")

identical(table1, table2)
```




Next Test:

Run Birdnet from R. Temporarily working on this in \_testCode.Rmd, then
move to newData.Rmd (#3)

-   Standardize info (lat/long, year round, etc). May want to rename
    birdNET_SelectionTable to include foldername info.

-   Create a datacheck for Birdnet per folder to ensure the values are
    as expected. Then rename using foldername info, and merge into a
    birdnet_params for the repo. Can I also add in the other user
    selected parameters? Also, for datacheck ensure there is a compiled
    text file (if not, then create it by compiling all the individual
    files)

-   birdNET_SelectionTable.txt- remove any rows where 'Begin Path'
    includes 'bad' in it (where it came from the 'bad' folder). Add
    measures to it using rRaven/WarbleR.

```{r birdnetCheck}
#remove any rows from the 'bad' folder

#


```
